Below are functions used to clean the data.

```{r}
source("00_requirements.R")

#######################################################
remove_na <- function(df) { ## Automatize the removal of missing observations.
  n_obs <- nrow(df)
  missing <- rep(FALSE, n_obs)
  for (obs_ind in 1:n_obs) {
    obs <- df[obs_ind, ]
    n_missing <- sum(is.na(obs))
    if(n_missing > 0) {
      missing[obs_ind] <- TRUE
    }
  }
  df_red <- df[!missing, ]

  return(df_red)
}
#######################################################
to_factors <- function(df) { ## Turn all character variables to factors.
  n_vars <- ncol(df)
  for (var_ind in 1:n_vars) {
    var <- df[, var_ind]
    if (class(var) == "character") {
      df[, var_ind] <- factor(var)
    }
  }
  return(df)
}
```

Load data

```{r}
FemaleLabor <- data.frame(read.csv("00_LaborForceData.csv"))
EducationExp <- data.frame(read.csv("00_GovEducationExpense.csv"))
BirthRate <- data.frame(read.csv("00_BirthRate.csv"))
FertilityRate <- data.frame(read.csv("00_FertilityRate.csv"))
ParentalLeave <- data.frame(read.csv("00_ParentalLeave.csv"))
TertiaryEnroll <- data.frame(read.csv("00_TertiaryEnrollmentRate.csv"))
```

We'll remove any unnecessary variables from the datasets then apply the functions defined above to each dataset.

```{r}
keep_col <- c("REF_AREA", "REF_AREA_LABEL", "TIME_PERIOD", "OBS_VALUE")

edu_clean <- EducationExp[,keep_col]
flabor_clean <- FemaleLabor[,keep_col]
brate_clean <- BirthRate[,keep_col]
fert_clean <- FertilityRate[,keep_col]
leave_clean <- ParentalLeave[,-c(1,2)]
enroll_clean <- TertiaryEnroll[,-c(1,2)]
```

```{r}
df_list <- list(edu_clean, flabor_clean, brate_clean, fert_clean, leave_clean, enroll_clean)

#check for missing values
for (d in df_list) {
  print(sum(is.na(d)))
}
```
Parental leave and Tertiary Enrollment are structured a bit differently from the other datasets. We need to separate male and female values and consider what we want to include in our model.

```{r}
#separating male and female values into their own columns
leave_sep <- leave_clean %>%
  pivot_wider(names_from = Disaggregation, values_from = Value, names_prefix = "days_")

enroll_sep <- enroll_clean %>%
  pivot_wider(names_from = Disaggregation, values_from = Value, names_prefix = "enrolled_")
```

We want to combine all of our data into one dataframe. However, some of the column names are not very descriptive so we must change the column names first.

```{r}
colnames(edu_clean)[colnames(edu_clean) == "OBS_VALUE"] <- "edu_exp"
colnames(flabor_clean)[colnames(flabor_clean) == "OBS_VALUE"] <- "female_labor_rate"
colnames(brate_clean)[colnames(brate_clean) == "OBS_VALUE"] <- "birth_rate"
colnames(fert_clean)[colnames(fert_clean) == "OBS_VALUE"] <- "fertility_rate"
```

Now we can merge all of our datasets into one dataframe

```{r}
df <- edu_clean %>%
  left_join(flabor_clean, by = c("REF_AREA", "REF_AREA_LABEL", "TIME_PERIOD")) %>%
  left_join(brate_clean, by = c("REF_AREA", "REF_AREA_LABEL", "TIME_PERIOD")) %>%
  left_join(fert_clean, by = c("REF_AREA", "REF_AREA_LABEL", "TIME_PERIOD")) %>%
  left_join(leave_sep, by = c("REF_AREA" = "Country.Code", "REF_AREA_LABEL" = "Country.Name", "TIME_PERIOD" = "Year")) %>%
  left_join(enroll_sep, by = c("REF_AREA" = "Country.Code", "REF_AREA_LABEL" = "Country.Name", "TIME_PERIOD" = "Year"))

sum(is.na(df))
colSums(is.na(df)) / nrow(df)
```










