---
title: "The Impacts of Female Labor"
output: pdf_document
date: "2025-11-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("12_dataSaving.RData")
```

# Abstract:

Using data from the World Bank, we ran multiple different regression models to explore the effects that education expenditure and fertility rates have on female labor force participation. By using other variables like parental leave and tertiary school enrollment as control variables we are able to reduce omitted variable bias. After running multiple different models we found that higher education expenditure associates with higher female labor force participation. Fertility rates however, displayed a non-linear, U-shaped relationship with female labor force participation. Understanding of these relationships can assist policy makers with improving labor market outcomes and overall economic performance.

# Introduction:
The purpose of this analysis is the examine the relationships between specific variables and the female labor participation rate. We focus mainly on how government expenditure on education affects the female labor participation rate and on how the fertility rate affects the female labor participation rate. We utilize regression modeling tools to measure the effect that these variables have on the female labor force participation rate. By exploring how the female labor force participation rate is affected numerous variables, we can understand how the female labor force participation rate affects the overall economy. Having this information is important for policy makers in both developing and developed countries. Increased female labor participation leads to lower poverty rates and more stable households. This information can be used to bring about economic stability and growth, better access to education, and gender equality.

# Data:

Our analysis uses global datasets from the World Bank on female labor participation rates, government expenditure on education, birth rates, fertility rates, parental leave, tertiary school enrollment, and GDP. After removing unnecessary columns and creating consistent column names, we merged all the datasets into one dataframe. We then sorted the countries by the amount of missing values they had and removed any country with over 25% missing values.

### Female Labor Participation Rate:

The female labor participation rate is represented by female_labor_rate in our data. It is a quantitative variable that measures the female labor force participation rate in terms of % of female population ages 15+. It contains data for almost all countries in the world and spans from 1990 to 2024. Data can be found at https://data360.worldbank.org/en/indicator/WB_WDI_SL_TLF_CACT_FE_ZS.

### Government Expenditure on Education

Government expenditure on education is represented by edu_exp in our data. It is a quantitative variable that measures the government expenditure on education in terms of total % of GDP. It contains data for almost all countries in the world and spans from 1970 to 2023. Data can be found at https://data360.worldbank.org/en/indicator/WB_WDI_SE_XPD_TOTL_GD_ZS.

### Birth Rate

Birth rate is represented by birth_rate in our data. It is a quantitative variable that measures the crude birth rate per 1000 people. It contains data globally and spans from 1960 to 2023. Data can be found at https://data360.worldbank.org/en/indicator/WB_WDI_SP_DYN_CBRT_IN.

### Fertility Rate

Fertility rate is represented by fertility_rate in our data. It is a quantitative variable that measures the total total fertility rates in terms of births per woman. It contains data globally and spans from 1960 to 2023. Data can be found at https://data360.worldbank.org/en/indicator/WB_WDI_SP_DYN_TFRT_IN.

### Maternal Leave

Maternal leave is represented by days_female in our data. It is a quantitative variable that measures the total length of paid maternal leave in terms of calendar days. It containts data globally and spans from 1970 to 2023. Data can be found at https://genderdata.worldbank.org/en/indicator/sh-par-leve?.

### Female Tertiary School Erollment

Female tertiary school enrollment is represented by enrolled_female in our data. It is a quantitative variable that measures female tertiary school enrollment in terms of % gross. It contains data globally and spans from 1970 to 2024. Data can be found at http://genderdata.worldbank.org/en/indicator/se-ter-enrrl.

### GDP

GDP is represented by gdp in our data. It is a quantitative variable that measures a country's GDP in terms of US Dollars. It contains data globally that spans from 1960 to 2024. Data can be found at https://data360.worldbank.org/en/indicator/WB_WDI_NY_GDP_MKTP_CD.

# Visualization:
Below is a histogram of Government Expenditure on Education (% of GDP).

```{r echo=FALSE}
summary(df$edu_exp)
hist(df$edu_exp, breaks = 75,
     main = "Distribution of Government Expenditure on Education",
     xlab = "Government Expenditure on Education (% of GDP)")
```
Below is a histogram of Birth Rate (per 1000 people).

```{r echo=FALSE}
summary(df$birth_rate)
hist(df$birth_rate, breaks = 50,
     main = "Distribution of Birth Rate",
     xlab = "Birth Rate (per 1000 people)")
```
Below is a histogram of Female Labor Participation Rate (% of female population age 15+).

```{r echo=FALSE}
summary(df$female_labor_rate)
hist(df$female_labor_rate, breaks = 50,
     main = "Distribution of Female Labor Participation Rate",
     xlab = "Female Labor Participation Rate (% of female population age 15+)")
```

We plan on using more than just histograms in our report but these are our current visualizations we plan on using.

# Analysis:

As seen above, we will use these graphs and future planned scatter/box plots and other necessary visuals, to cross reference this data and find a conclusion on our two questions we will solve. We'll test Birthrates against female labor participation across different countries and review the data. We'll then pit it against other forms of data that could contribute to female labor to see if another factor could be the main cause. We'll do the same process with government spending on education and provide all findings we come across.