---
title: "The Impacts of Female Labor"
output: pdf_document
date: "2025-11-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract:

This study investigates the what impact female labor has on a countries birthrate. Additionally, we also want to explore if how much a country's expenditure on education affects the total amount of female contribution in the workforce. We aim to use visualizations such as histograms, box plots, and scatter plots to answer this question. Using these visualizations will allow us to see the distribution of key variables to help give a clear answer.

# Introduction:

We wanted to know if countries with higher female labor contributed to lower birthrates in that country. Our hypothesis is that we believe that female labor does decrease birthrates. This analysis also explores if government spending on education and schools assist in increasing female labor participation. If there is a significant correlation, countries with declining populations could use this information to balance employment, education, and growth to stable levels. 

# Data:



# Visualization:

```{r echo=FALSE}

```

# Analysis: